Variational Monte Carlo on bosonic systems

### 2018-01-22 project_1/article/article.tex
Add trial wavefunction
Add motivation to finding the local energy

\section{Theory}
    To model the trapped bosonic gas particles we use the potential
    \begin{align}
        V_{\text{ext}}(\vf{r})
        &=
        \begin{cases}
            \half m\omega^2r^2 & (\text{S}), \\
            \half m [
                \omega^2(x^2 + y^2) + \omega_z^2z^2
            ] & (\text{E}),
        \end{cases}
    \end{align}
    where we can choose between a spherical (S) or an elliptical (E) harmonic
    trap. The two-body Hamiltonian of the system is given by
    \begin{align}
        H = \sum_{i}^{N}h(\vf{r}_i) + \sum_{i < j}^{N}w(\vf{r}_i, \vf{r}_j),
    \end{align}
    where the single particle one body operator, $h$, is given by
    \begin{align}
        h(\vf{r}_i) = -\frac{\hbar^2}{2m}\nabla_i^2
        + V_{\text{ext}}(\vf{r}_i),
    \end{align}
    and the two-body interaction operator, $w$, is
    \begin{align}
        w(\vf{r}_i, \vf{r}_j)
        = \begin{cases}
            \infty & |\vf{r}_i - \vf{r}_j| \leq a, \\
            0 & |\vf{r}_i - \vf{r}_j| > a,
        \end{cases}
    \end{align}
    where $a$ is the hard sphere of the particle.

>>>>

\section{Theory}
    To model the trapped bosonic gas particles we use the potential
    \begin{align}
        V_{\text{ext}}(\vf{r})
        &=
        \begin{cases}
            \half m\omega^2r^2 & (\text{S}), \\
            \half m \bigl[
                \omega^2(x^2 + y^2) + \omega_z^2z^2
            \bigr] & (\text{E}),
        \end{cases}
    \end{align}
    where we can choose between a spherical (S) or an elliptical (E) harmonic
    trap. The two-body Hamiltonian of the system is given by
    \begin{align}
        H = \sum_{i}^{N}h(\vf{r}_i) + \sum_{i < j}^{N}w(\vf{r}_i, \vf{r}_j),
    \end{align}
    where the single particle one body operator, $h$, is given by
    \begin{align}
        h(\vf{r}_i) = -\frac{\hbar^2}{2m}\nabla_i^2
        + V_{\text{ext}}(\vf{r}_i),
    \end{align}
    and the two-body interaction operator, $w$, is
    \begin{align}
        w(\vf{r}_i, \vf{r}_j)
        = \begin{cases}
            \infty & |\vf{r}_i - \vf{r}_j| \leq a, \\
            0 & |\vf{r}_i - \vf{r}_j| > a,
        \end{cases}
    \end{align}
    where $a$ is the hard sphere of the particle. The trial wavefunction we will
    be looking at is given by
    \begin{align}
        \Psi_T(\vf{r})
        &= \bra{\vf{r}}\ket{\Psi_T}
        = \prod_i^N g(\alpha, \beta, \vf{r}_i)
        \prod_{i < j}^N f(a, \vf{r}_i, \vf{r}_j),
    \end{align}
    where $\alpha$ and $\beta$ are variational parameters and
    \begin{align}
        \vf{r} = (\vf{r}_1, \vf{r}_2, \dots, \vf{r}_N, \alpha, \beta).
    \end{align}
    Here $g$ are the single-particle wavefunctions given by
    \begin{align}
        g(\alpha, \beta, \vf{r}_i)
        = \exp\bigl[
            -\alpha(x_i^2 + y_i^2 + \beta z_i^2)
        \bigr] \equiv \phi(\vf{r}_i),
    \end{align}
    and the correlation wavefunction, $f$, is
    \begin{align}
        f(a, \vf{r}_i, \vf{r}_j)
        &=
        \begin{cases}
            0 & |\vf{r}_i - \vf{r}_j| \leq a, \\
            \Bigl(
                1 - \frac{a}{|\vf{r}_i - \vf{r}_j|}
            \Bigr) & |\vf{r}_i - \vf{r}_j| > a.
        \end{cases}
    \end{align}

    \subsection{Local energy}
        As the many-body wavefunction creates a very large configuration space,
        where much of the wavefunction is small, we use the Metropolis algorithm
        in order to move towards regions in configuration space with
        ``sensible'' values. We define the \textit{local energy}, $E_L{\vf{r}}$,
        by
        \begin{align}
            E_L(\vf{r})
            &= \frac{\bra{\vf{r}}H\ket{\Psi_T}}{\bra{\vf{r}}\ket{\Psi_T}}.
        \end{align}
        If $\ket{\Psi_T}$ is an exact eigenfunction of the Hamiltonian, $E_L$
        will be constant. The closer $\ket{\Psi_T}$ is to the exact wave
        function, the less variation in $E_L$ as a function of $\vf{r}$ we get.
        One of the most computationally intensive part of the VMC algorithm will
        be to compute $E_L$. We therefore find an analytical expression for
        $E_L$ in terms of the trial wavefunction.
